---
title: "Religion, Cultural Norms, and Wellbeing: A Bayesian Multilevel Analysis"
author: Scott Claessens, Kyle Fischer, Guy Lavender Forsyth, Thanos Kyritsis, and Quentin Atkinson
date: "`r format(Sys.Date())`"
output: github_document
bibliography: r-references.bib
csl: nature.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(dev = 'svg')
```

```{r echo=F}
loadd(dLong, scree, looCompare1, looCompare2, 
      summary1.2, summary1.3, cond1.2, cond1.3,
      f1.2.2, f1.2.3, f1.2.4, f1.3.2, f1.3.3, f1.3.4)
```

# Analytic Strategy

Do religious people report higher wellbeing? And is this relation moderated by perceived cultural norms of religion? In order to answer these questions, we constructed a series of Bayesian multilevel cumulative link models, which are appropriate for highly nested ordinal data, and fitted these models to the data in stacked format, with separate rows for each individual wellbeing item (n = `r nrow(dLong)`). The key dependent variable for all our models was perceived wellbeing, with wellbeing subscales and items included as random effects. The main predictor, religiosity, was defined as the first principal component in a principal component analysis including all nine religiosity items, which explained about 80% of the variance (see scree plot in Figure 1).

```{r scree, echo=FALSE, warning=FALSE, message=FALSE}
scree
```

<span style="font-size:0.9em">***Figure 1.*** _Scree plot for principal components analysis of nine religion items in the dataset. Following both the point of inflexion and Kaiser's criterion[@Kaiser1960] (i.e. retaining principal components with eigenvalues above one), we should retain one principal component._</span><br><br>

The second predictor, perceived cultural norms of religion, was the average of two items measuring the extent to which participants perceive religion being important in their country. Our models also included a number of control variables (age, gender, and scaled GDP), and random intercepts and slopes grouped by wellbeing subscale and item, participant ID, participant country, and participant religious denomination. Finally, in another set of models, we additionally let country-level random intercepts covary according to their linguistic proximity (see below for more details).

We introduced these variables into our models incrementally. Initially, we ran an intercept-only model (the outcome variable predicted by demographic controls and random intercepts, without religiosity and cultural norms) to determine a baseline effect. Next, we added religiosity as a main predictor (research question 1), and then cultural norms and its interaction with religiosity (research question 2). Lastly, we repeated these analyses by adding a linguistic proximity matrix between the 24 countries in the sample. The full model was as follows:

```{r eval=FALSE}
wellbeing ~ 1 + religion*cnorm + (1 + religion*cnorm | subscale/item) + 
  (1 + religion*cnorm | iso) + (1 + religion*cnorm | denomination) + 
  (1 | subject) + age + gender + gdp_scaled
```

where `wellbeing`, `religion`, and `cnorm` correspond to the three main variables as described above, `subscale` and `item` refer to the wellbeing subscale and item, `iso` is the ISO-3166-1 code assigned to the participant's country, `denomination` is the participant's reported religious denomination, `subject` is the participant's ID, and `age`, `gender`, and `gdp_scaled` are the demographic variables.

The final step of accounting for linguistic proximity was necessary in order to control for Galton's problem[@Naroll1961], i.e. observed similarities between closely-related cultures being due to their common history. To produce this matrix, we joined all language family trees published in Glottolog v3.0[@Glottolog] into an undated and unresolved global language tree. We then calculated proximity <img src="https://render.githubusercontent.com/render/math?math=s"> between two languages <img src="https://render.githubusercontent.com/render/math?math=j"> and <img src="https://render.githubusercontent.com/render/math?math=k"> as the distance (in number of nodes traversed) of their most recent common ancestor <img src="https://render.githubusercontent.com/render/math?math=i"> to the root of the tree, through the formula:

<p align="center"><img src="https://render.githubusercontent.com/render/math?math=%5CLARGE%20s_{jk}%20=%20%5Cfrac{n_{r}-n_{i}}{n_{r}}"></p>

where <img src="https://render.githubusercontent.com/render/math?math=n_{r}"> is the maximum path length (in number of nodes traversed) leading to the pan-human root <img src="https://render.githubusercontent.com/render/math?math=r">, and <img src="https://render.githubusercontent.com/render/math?math=n_{i}"> is the maximum path length leading to node  <img src="https://render.githubusercontent.com/render/math?math=i">.

We then paired this data with speaker percentages from Ethnologue 21[@Ethnologue], and set linguistic proximity between two countries as the cultural proximities between all languages spoken within them, weighted by respective speaker percentages (1 per mille was used as a realistic cut-off point), through the formula:

<p align="center"><img src="https://render.githubusercontent.com/render/math?math=%5CLARGE%20w_{lm}%20=%20{%5CSigma}{%5CSigma}p_{lj}p_{mk}s_{jk}"></p>

where <img src="https://render.githubusercontent.com/render/math?math=p_{lj}"> is the percentage of the population in nation <img src="https://render.githubusercontent.com/render/math?math=l"> speaking language <img src="https://render.githubusercontent.com/render/math?math=j">, <img src="https://render.githubusercontent.com/render/math?math=p_{mk}"> is the percentage of the population in nation <img src="https://render.githubusercontent.com/render/math?math=m"> speaking language <img src="https://render.githubusercontent.com/render/math?math=k">, and <img src="https://render.githubusercontent.com/render/math?math=s_{jk}"> is the proximity measure between languages <img src="https://render.githubusercontent.com/render/math?math=j"> and <img src="https://render.githubusercontent.com/render/math?math=k">[@Claessens2020;@Eff2008].

Bayesian model priors were based on prior predictive simulation. All models had zero divergent transitions and converged normally (<img src="https://render.githubusercontent.com/render/math?math=%5Chat{R}"> < 1.05). The analysis pipeline was run in R v4.0.2.[@RCoreTeam] using the _drake_ package[@Landau2018] for reproducible high performance computing, the _brms_ package[@Burkner2017] for model fitting, and the _loo_ package[@Vehtari2017] for model comparison with approximate leave-one-out cross-validation using subsampling[@Magnusson2019].

# Results

## Research Question 1

Does religiosity predict wellbeing? Yes. The model including the religiosity predictor fits the data better than the null model (difference in ELPD = `r round(looCompare1[2,1], 1)`, SE = `r round(looCompare1[2,2], 1)`) and the fixed effect shows that religiosity is associated with greater overall wellbeing (b = `r round(summary1.2$fixed["religion","Estimate"], 2)`, SE = `r round(summary1.2$fixed["religion","Est.Error"], 2)`, 95% credible interval [`r round(summary1.2$fixed["religion","l-95% CI"], 2)` `r round(summary1.2$fixed["religion","u-95% CI"], 2)`]; Figure 2).

```{r cond1.2, echo=FALSE, warning=FALSE, message=FALSE}
cond1.2
```

<span style="font-size:0.9em">***Figure 2.*** _The effect of religiosity on overall wellbeing. Regression lines and 95% credible interval shaded areas are predictions from our Bayesian multilevel cumulative link model, controlling for age, gender, and GDP but not controlling for linguistic proximity between countries._</span><br><br>

This positive effect of religiosity holds for social, psychological, and physical wellbeing subscales, and most individual wellbeing items (the effect is particularly pronounced for meaningfulness), though the 95% credible intervals include zero for mobility, negative affect, physical pain, and medical dependence (Figure 3). The positive effect of religiosity on wellbeing also holds for most countries in the sample, but the 95% credible intervals overlap with zero for Singapore, the Netherlands, the UK, Spain, Denmark, Chile, and Canada (Figure 4). The positive effect of religiosity also holds across most religious denominations in the sample, but the 95% credible intervals include zero for Shinto, Azhari and Alevi Muslim, Jain, Evangelical Christian, Christian (denomination unspecified), and African religions (Figure 5).

```{r f1.2.2, echo=FALSE, warning=FALSE, message=FALSE}
f1.2.2
```

<span style="font-size:0.9em">***Figure 3.*** _Density plot showing how the effect of religiosity on wellbeing varies across individual wellbeing items. The visualised posterior distribution is from our Bayesian multilevel cumulative link model, controlling for age, gender, and GDP but not controlling for linguistic proximity between countries. Densities are posterior distributions for random effects, points are posterior medians, lines are 66% and 95% credible intervals._</span><br><br>

```{r f1.2.3, echo=FALSE, warning=FALSE, message=FALSE}
f1.2.3
```

<span style="font-size:0.9em">***Figure 4.*** _Density plot showing how the effect of religiosity on wellbeing varies across individual countries. The visualised posterior distribution is from our Bayesian multilevel cumulative link model, controlling for age, gender, and GDP but not controlling for linguistic proximity between countries. Densities are posterior distributions for random effects, points are posterior medians, lines are 66% and 95% credible intervals._</span><br><br>

```{r f1.2.4, echo=FALSE, warning=FALSE, message=FALSE}
f1.2.4
```

<span style="font-size:0.9em">***Figure 5.*** _Density plot showing how the effect of religiosity on wellbeing varies across individual religious denominations. The visualised posterior distribution is from our Bayesian multilevel cumulative link model, controlling for age, gender, and GDP but not controlling for linguistic proximity between countries. Densities are posterior distributions for random effects, points are posterior medians, lines are 66% and 95% credible intervals._</span><br><br>

Do these results change when controlling for cultural connections between countries (i.e., linguistic proximity)? Essentially, no. However, in this model, the 95% credible intervals no longer include zero for some countries (the UK and Canada) and some religious denominations (Shinto and African religions), revealing a positive effect of religion for these clusters when controlling for linguistic proximity.

## Research Question 2

Does the effect of religiosity on wellbeing depend on perceiving others to be religious? Our results indicate that the effect of religiosity on wellbeing is only weakly and inconsistently moderated by religious norms. A model including an interaction between religiosity and religious norms does fit the data better than one which excludes religious norms (difference in ELPD = `r round(looCompare2[2,1], 1)`, SE = `r round(looCompare2[2,2], 1)`; cautious interpretation is needed here due to substantial subsampling error, SSE = `r format(round(looCompare2[2,3], 1), nsmall = 1)`). The interaction fixed effect from this model indicates that the effect of religiosity on overall wellbeing is minimally increased when participants perceive religion to be important in their country (b = `r format(round(summary1.3$fixed["religion:cnorm","Estimate"], 2), nsmall = 2)`, SE = `r round(summary1.3$fixed["religion:cnorm","Est.Error"], 2)`, 95% credible interval [`r round(summary1.3$fixed["religion:cnorm","l-95% CI"], 2)` `r round(summary1.3$fixed["religion:cnorm","u-95% CI"], 2)`]; Figure 6), though the 95% credible interval for this effect includes zero.

```{r cond1.3, echo=FALSE, warning=FALSE, message=FALSE}
cond1.3
```

<span style="font-size:0.9em">***Figure 6.*** _The interaction between religiosity and cultural norms. Regression lines and 95% credible interval shaded areas are predictions from our Bayesian multilevel cumulative link model, controlling for age, gender, and GDP but not controlling for linguistic proximity between countries._</span><br><br>

Why only this small and inconsistent interaction effect? Our modelling indicates that it is likely due to variation in the interaction effect across individual wellbeing items and individual countries. How does the interaction effect fare across individual wellbeing items? Figure 7 supports the hypothesised positive interaction effect for most individual wellbeing items, including general satisfaction with health, energy for everyday life, satisfaction with sleep, ability to perform daily living activities, ability to concentrate, acceptance of bodily appearance, satisfaction with oneself, frequency of negative feelings, satisfaction with personal relationships, and satisfaction with sex life. However, we find that the interaction effect is negative for wellbeing items capturing pain, medical dependence, and mobility, and for others the 95% credible interval includes zero.

```{r f1.3.2, echo=FALSE, warning=FALSE, message=FALSE}
f1.3.2
```

<span style="font-size:0.9em">***Figure 7.*** _Density plot showing how the interaction effect varies across individual wellbeing items. The visualised posterior distribution is from our Bayesian multilevel cumulative link model, controlling for age, gender, and GDP but not controlling for linguistic proximity between countries. Densities are posterior distributions for random effects, points are posterior medians, lines are 66% and 95% credible intervals._</span><br><br>

How does the interaction effect vary across individual countries and religious denominations? Only the USA, India, and Australia show positive interaction effects with 95% credible intervals above zero (Figure 8). Across individual religious denominations, all 95% credible intervals include zero without large differences between denominations (Figure 9).

```{r f1.3.3, echo=FALSE, warning=FALSE, message=FALSE}
f1.3.3
```

<span style="font-size:0.9em">***Figure 8.*** _Density plot showing how the interaction effect varies across individual countries. The visualised posterior distribution is from our Bayesian multilevel cumulative link model, controlling for age, gender, and GDP but not controlling for linguistic proximity between countries. Densities are posterior distributions for random effects, points are posterior medians, lines are 66% and 95% credible intervals._</span><br><br>

```{r f1.3.4, echo=FALSE, warning=FALSE, message=FALSE}
f1.3.4
```

<span style="font-size:0.9em">***Figure 9.*** _Density plot showing how the interaction effect varies across individual religious denominations. The visualised posterior distribution is from our Bayesian multilevel cumulative link model, controlling for age, gender, and GDP but not controlling for linguistic proximity between countries. Densities are posterior distributions for random effects, points are posterior medians, lines are 66% and 95% credible intervals._</span><br><br>

Most of these results are unchanged when controlling for linguistic proximity between countries, except that in this model the 95% credible intervals for the interaction effect include zero for some wellbeing items (mobility and personal relations).

# References
