---
title: "Religion and Wellbeing - Blinded Analysis"
author: Scott Claessens
date: "`r format(Sys.Date())`"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r options, echo=F}
options(width = 200)
```

The purpose of this document is to report the results of a blinded analysis relating religiosity to wellbeing.

_NOTE: this is dummy code that only uses a small slice of the dataset to speed up computation._

# Initial religion PCA and data organisation

To find a single predictor for our models that captures "religiosity", we first run a principal components analysis with all nine of the religion items to see how many factors it produces.

A scree plot of this PCA reveals that a single factor is enough.

```{r scree, echo=F}
readd(scree)
```

For modelling, we then (1) add this principal component to the dataset as `religion`, (2) create a composite variable of `cnorm_1` and `cnorm_2` called `cnorm`, and (3) organise the data in long format. The data then look like this (only the top 20 rows):

```{r data, echo=F}
readd(dLong) %>%
  select(wellbeing, item, subscale, religion, cnorm, 
         subject, country, denomination, age, gender, gdp_scaled) %>%
  head(n = 20) %>%
  kable() %>%
  kable_styling() %>%
  scroll_box(width = "100%")
```

# Models without controlling for spatial autocorrelation for countries

## Question 1: Does religiosity predict wellbeing?

First, we fit a "null" random-intercept-only ordinal multilevel model. This model includes demographic control variables (age, gender, and GDP) and random intercepts for wellbeing items nested within wellbeing subscales, and for participants within countries and denominations: `wellbeing ~ 1 + (1 | subscale/item) + (1 | iso) + (1 | denomination) + (1 | subject) + age + gender + gdp_scaled`.

Here's the summary for this model.

```{r summary1.1, echo=F}
readd(summary1.1)
```

We get cutpoints for the ordinal intercept, main effects for the demographic control variables, and standard deviations from the random effects.

Next, we add a fixed effect for the religiosity predictor, as well as (1) a random slope grouped by wellbeing item nested within subscale, (2) a random slope grouped by country, and (3) a random slope grouped by denomination: `wellbeing ~ 1 + religion + (1 + religion | subscale/item) + (1 + religion | iso) + (1 + religion | denomination) + (1 | subject) + age + gender + gdp_scaled`. This gives us a different effect of religion for every wellbeing subscale, every wellbeing item, every country, and every religious denomination.

Does this second model improve the fit? We would expect it to if `religion` is a good predictor of wellbeing.

```{r looCompare1, echo=F}
readd(looCompare1)
```

What's the summary for this model?

```{r summary1.2, echo=F}
readd(summary1.2)
```

The fixed effect of `religion` is of interest.

Let's visualise the effect of religiosity on _overall_ wellbeing.

```{r cond1.2, echo=F}
readd(cond1.2)
```

We can also split this effect into separate wellbeing subscales...

```{r forest1.2.1, echo=F}
readd(f1.2.1)
```

... and by individual wellbeing items...

```{r forest1.2.2, echo=F}
readd(f1.2.2)
```

... and by country...

```{r forest1.2.3, echo=F}
readd(f1.2.3)
```

... and by religious denomination.

```{r forest1.2.4, echo=F}
readd(f1.2.4)
```

## Question 2: Is there an interaction with perceived cultural norms?

We then add a fixed interaction effect with `cnorm`, and allow this interaction effect to vary as a random slope over wellbeing items nested within subscales, as a random slope over countries, and as a random slope over denominations: `wellbeing ~ 1 + religion*cnorm + (1 + religion*cnorm | subscale/item) + (1 + religion*cnorm | iso) + (1 + religion*cnorm | denomination) + (1 | subject) + age + gender + gdp_scaled`. In this model, every wellbeing subscale, every individual wellbeing item, every country, and every denomination has its own interaction between religion and cultural norms.

Does this model improve the fit over the previous model? If there's an interaction effect, we would expect it to.

```{r looCompare2, echo=F}
readd(looCompare2)
```

Here's the summary for this model.

```{r summary1.3, echo=F}
readd(summary1.3)
```

The fixed effect `religion:cnorm` is of interest.

Plot the interaction.

```{r cond1.3, echo=F}
readd(cond1.3)
```

We can also get forest plots of the interaction parameter for wellbeing subscales...

```{r forest1.3.1, echo=F}
readd(f1.3.1)
```

... for individual wellbeing items...

```{r forest1.3.2, echo=F}
readd(f1.3.2)
```

... for countries...

```{r forest1.3.3, echo=F}
readd(f1.3.3)
```

... and for religious denominations.

```{r forest1.3.4, echo=F}
readd(f1.3.4)
```

# Models controlling for spatial autocorrelation for countries

Do our results change when we control for spatial autocorrelation among countries? Some countries may have similar wellbeing scores simply because they are closer together.

## Question 1: Does religiosity predict wellbeing?

First, we fit a "null" random-intercept-only ordinal multilevel model, controlling for spatial autocorrelation for country random intercepts: `wellbeing ~ 1 + (1 | subscale/item) + (1 | gr(iso, cov = geo) + (1 | denomination) + (1 | subject) + age + gender + gdp_scaled`.

Here's the summary for this model.

```{r summary2.1, echo=F}
readd(summary2.1)
```

Next, we add the religiosity predictor: `wellbeing ~ 1 + religion + (1 + religion | subscale/item) + (1 | gr(iso, cov = geo, id = "iso)) + (0 + religion | gr(iso, id = "iso")) + (1 + religion | denomination) + (1 | subject) + age + gender + gdp_scaled`. We only allow the random intercept for countries (not the random slope) to covary according to geographic distance, but still model the random intercept and random slope as correlated.

Does this second model improve the fit? We would expect it to if `religion` is a good predictor of wellbeing.

```{r looCompare3, echo=F}
readd(looCompare3)
```

What's the summary for this model?

```{r summary2.2, echo=F}
readd(summary2.2)
```

The fixed effect of `religion` is of interest.

Let's visualise the effect of religiosity on _overall_ wellbeing, controlling for spatial autocorrelation.

```{r cond2.2, echo=F}
readd(cond2.2)
```

We can also split this effect into separate wellbeing subscales...

```{r forest2.2.1, echo=F}
readd(f2.2.1)
```

... and by individual wellbeing items...

```{r forest2.2.2, echo=F}
readd(f2.2.2)
```

... and by country...

```{r forest2.2.3, echo=F}
readd(f2.2.3)
```

... and by religious denomination.

```{r forest2.2.4, echo=F}
readd(f2.2.4)
```

## Question 2: Is there an interaction with perceived cultural norms?

We then add an interaction effect with `cnorm`: `wellbeing ~ 1 + religion*cnorm + (1 + religion*cnorm | subscale/item) + (1 | gr(iso, cov = geo, id = "iso)) + (0 + religion*cnorm | gr(iso, id = "iso")) + (1 + religion*cnorm | denomination) + (1 | subject) + age + gender + gdp_scaled`. In this model, every wellbeing subscale, every individual wellbeing item, every country, and every denomination has its own interaction between religion and cultural norms.

Does this model improve the fit over the previous model? If there's an interaction effect, we would expect it to.

```{r looCompare4, echo=F}
readd(looCompare4)
```

Here's the summary for this model.

```{r summary2.3, echo=F}
readd(summary2.3)
```

The fixed effect `religion:cnorm` is of interest.

Plot the interaction.

```{r cond2.3, echo=F}
readd(cond2.3)
```

We can also get forest plots of the interaction parameter for wellbeing subscales...

```{r forest2.3.1, echo=F}
readd(f2.3.1)
```

... for individual wellbeing items...

```{r forest2.3.2, echo=F}
readd(f2.3.2)
```

... for countries...

```{r forest2.3.3, echo=F}
readd(f2.3.3)
```

... and for religious denominations.

```{r forest2.3.4, echo=F}
readd(f2.3.4)
```

# Session Info

```{r sessionInfo}
sessionInfo()
```
